---
title: "Conditionals Loops"
author: "Ruth Enriquez"
date: "2022-08-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1

activity taken from: <https://allisonhorst.github.io/EDS_221_programming-essentials/activities/day_3_activities.html>

## Task 1

-   Create an object called pm2_5 with a value of 48

-   Write an `if - else if - else` statement that returns "Low to moderate risk" if `pm2_5` (for Particulate Matter 2.5) is less than 100, "Unhealthy for sensitive groups" if PM 2.5 is 100 \<= pm2_5 \< 150, and "Health risk present" if PM 2.5 is \>= 150.

-   Test by changing the value of your pm2_5 object and re-running your statement to check.

```{r}
pm2_5 <- 48

if (pm2_5 < 100) {
  print("Low to moderate risk")
} else if (pm2_5 >= 100 & pm2_5 < 150) {
  print("Unhealthy for sensitive groups")
} else {
  print("Health risk present")
}
```

## Task 2

-   Store the string "blue whale" as an object called `species`. Write an if statement that returns "You found a whale!" if the string "whale" is detected in species, otherwise return nothing. Test by changing the species string & re-running to see output.

```{r}
species <- "blue whale"

if (str_detect(species, "whale")) {
  print("You found a whale!")
}
```

Task 3

-   Create a vector stored as max_airtemp_c with a value of 24.1. Write an if else statement that will print "Temperature too high" if max_airtemp_c is greater than 27, or "Temperature OK" if temperature is less than or equal to 27.

```{r}
max_airtemp_c <-24.1
if (max_airtemp_c > 27) {
  print("Temperature too high")
} else if (max_airtemp_c <= 27) {
  print("Temperature OK")
} 
```

## Task 4

-   Store the base price of a burrito as `base_burrito` with a value of 6.50. Store `main_ingredient` with a starting string of "veggie." Write a statement that will return the price of a burrito based on what a user specifies as "main_ingredient" (either "veggie", "chicken" or "steak") given the following:

    -   A veggie burrito is the cost of a base burrito

    -   A chicken burrito costs 3.00 more than a base burrito

    -   A steak burrito costs 3.25 more than a base burrito

```{r}
base_burrito <- 6.5
main_ingredint <- "veggie"

if (main_ingredint =="veggie") {
  print(base_burrito)
} else if (main_ingredint == "chicken") {
  print(base_burrito + 3.00)
} else if (main_ingredint == "steak") {
  print(base_burrito + 3.25)
}
```

For Loops

Task 5

-   Create a new vector called `fish` that contains the values `8, 10, 12, 23` representing counts of different fish types in a fish tank (goldfish, tetras, guppies, and mollies, respectively). Write a for loop that iterates through `fish`, and returns what proportion of total fish in the tank are that species. Assume that these counts represent all fish in the tank.

```{r}
fish <- c(8, 10, 12, 23)
fish_species <- c("goldfish", "tetra", "guppies", "mollies")

for (i in seq_along(fish)) {
  if (fish_species[i] == "goldfish") {
    count =(fish[i]/sum(fish)) * 100
  } else if (fish_species[i] == "tetra") {
    count =(fish[i]/sum(fish)) * 100
  } else if (fish_species[i] == "guppies") {
    count =(fish[i]/sum(fish)) * 100
  } else if (fish_species[i] == "mollies") {
    count =(fish[i]/sum(fish)) * 100
  }
 print(count)
}
```

## Task 6

-   There is an existing vector in R called month.name that contains all month names (just try running month.name in the Console to check it out). Write a for loop that iterates over all months in month.name and prints "January is month 1," "February is month 2", etc.

-   Hint: you can index values in the month.name vector just like you would any other vector (e.g., try running month.name[5]).

```{r}
for (i in seq_along(month.name)) {
     print(paste(month.name[i], "is month", i))
  }
```

Part 2

